Systematic Models for Inductive Verification of Fault-Tolerant Distributed Systems

### 2016-11-10 doc/om-paper/paper.tex
paper wibbles

\title{Systematic Models for $k$-induction and Synthesis of Fault-Tolerant Distributed Systems}
\author{Ben and Lee}

@@

\lee{needs to be redone.}
Fault-tolerant distributed systems are the backbone of reliable life-critical systems, like avionics. Consequently, this class of systems demand high-assurance of correct design and implementation. Formal verification can be used to prove the correctness of their designs.

There is substantial literature on verifying fault-tolerant distributed systems~\cite{}. Most of this work uses interactive theorem-proving~\cite{}. Interactive theorem-proving is useful in the domain because of the complexity of the models and because distributed system verification is typically parameterized by the number of nodes; automated methods perform poorly along these two metrics. However, two problems with interactive verification are that it requires specialized expertise and it is labor intensive.

In contrast, model-checking is automated but has limited scalability. To address scalability, model-specific abstractions are employed, both of can lead to problems.

For example, verification abstractions are often ad-hoc and system-specific. For example, to model a process that can broadcast the values $\{0, 1\}$, or no value when faulty, one might define an enumerated type of broadcast values sent such as

$$\{0, 1, NONE\}$$

However, the enumeration conflates environmental artifacts (i.e., faults) and system artifacts (i.e., the non-faulty broadcast values). Conflating them makes formal model-to-code correspondence more difficult to achieve.

Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided. While these optimizations might be safe under particular assumptions, they can lead to unsound models when a system or its environment is modified.

@@

@@

Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs). ASMs are a generalization of finite state machines originally developed by Gurevich to generalize the Church-Turing thesis for abstract data structures~\cite{}. Since then, ASMs have become an engineering technique for simulating systems at different levels of abstraction.

\lee{http://www2.informatik.hu-berlin.de/sam/preprint/reisig203.pdf}
More precisely, we implement a sequential small-step ASM~\cite{}.
\begin{definition}[Abstract state machine]
Let $\mathcal{M} = (S, I, \rightarrow), \mathcal{M}' = (S', I', \rightarrow')$ be transition systems. Let $B \subseteq 2^S$ be a set of relations on $S$. We say that $\mathcal{M'}$ is an abstract state machine for $\mathcal{M} | \mathcal{M'}$ iff $S \cap S' = \emptyset$, and each transition $t \in \rightarrow'$ is of the form
$$\textnormal{if } b_0(s_0) \textnormal{ then } s'_0 \textnormal{ else if } b_1(s_1) \textnormal{ then } s'_1 \ldots \textnormal{ else } b_n(s_n) \textnormal{ then } s'_n$$
\end{definition}

@@

% ------------------------------------------------------------
\section{Extended Example: OM(1)}\label{sec:byz}

>>>>

\title{Systematic Models for Inductive Verification of Fault-Tolerant Distributed Systems}
\author{Ben and Lee}

@@

\lee{needs to be redone.}
Fault-tolerant distributed systems are famously complex to design and build, yet are the  backbone of life-critical systems. Consequently, this class of systems demand high-assurance of correct design and implementation. Formal verification can be used to prove the correctness of their designs.

There is substantial literature on verifying fault-tolerant distributed systems. Most of this work uses interactive theorem-proving~\cite{}. Interactive theorem-proving is useful in the domain because of the complexity of the models and because distributed system verification is typically parameterized by the number of nodes; automated methods perform poorly along these two metrics. However, two problems with interactive verification are that it requires specialized expertise and it is labor intensive.

In contrast, model-checking for this class of systems is automated~\cite{}, but has limited scalability, particularly in the context of modeling the nondetermism of faults and distributed computation. The way that scalability is usually addressed is by specialization and ad-hoc model-specific abstractions. Specialization includes fixing a (small number) of nodes and topology for distributed protocols that are parameterized by the number of nodes and network.

Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value. To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$. However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values). (2) Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided. While these optimizations might be safe under particular assumptions, they require careful meta-level reasoning and justification and divorce the model from any realization of it.

However, recent advances in model-checking allow for better scalability and model fidelity, and they begin to blur the lines between model-checking and interactive theorem-proving. In particular, $k$-induction for infinite-state systems allows the specifier to assert $k$-inductive lemmas

  (We discuss the applicability of property-directed reachability in Section~\ref{sec:related}.)

@@

\lee{Talk about how to define an MFA in the environment}

@@

Our solution to the problem of constructing a mode-based state machine uses abstract state machines (ASMs). ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}.

In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:

\begin{definition}[Abstract state machine (ASM)]
Let $\mathcal{M} = (S, I, \rightarrow)$ be a transition system and let $B \subseteq 2^S$ be a set of predicates over $S$. Then $\mathcal{M}$ is an ASM iff the transition relation $\rightarrow$ is of the form
$$\textnormal{if } b_0(s_0) \textnormal{ then } s'_0 \textnormal{ else if } b_1(s_1) \textnormal{ then } s'_1 \ldots \textnormal{ else } b_n(s_n) \textnormal{ then } s'_n$$
\noindent
where $b_0, b_1, \ldots, b_n$ are arbitrary Boolean combinations of the predicates in $B$.
\end{definition}

Our use of ASMs to improve traceability is even more restrictive. We construct an ASM $\mathcal{M}$ that is synchronously composed with the system specification $\mathcal{S}$. Each state of the ASM represents a ``configurations'' of $S$ and is associated with some predicate over the states of $\mathcal{S}$.

@@

% ------------------------------------------------------------
\section{Putting it Together: Compositional Modeling and Verification}\label{sec:byz}
